Bayesian neural networks for sparse coding

### 2018-07-30 review_nips_2018/response_to_reviews_nips.tex

\documentclass{article}
\usepackage{nips_2018}
\usepackage{color}
\usepackage{mathtools}
\begin{document}

\title{NIPS 2018 Reply to Reviewers}
\date{}
\maketitle

\section*{Reviewer 1}
Q: What is the motivation for being Bayesian about the LISTA algorithm. How does in the uncertainty in the Bayesian LISTA procedure relate to the original linear model.

A: The LISTA algorithm introduces the weights in the model. Their point estimates are being inferred during the training phase, this leads to the uncertainty of the estimation. In our Bayesian model we infer the posterior distributions over these weights to estimate this uncertainty.
\newline
\newline

Q: how does the sparse promoting prior interact with the sparsity promoting non-linearity.

A: In the LISTA model the sequential estimates of beta are obtained at each layer using the previous estimate. At every layer the estimate becomes more sparse due to soft-thresholding non-linearity. We do not specify sparsity priors over the beta_l - we show that show that the distribution of these estimates can be approximated with spike-and-slab distribution, i.e. it sparsity naturally arises from Gaussian prior on NN weights and soft thresholding nonlinearity.
\newline
\newline

Q: The approximations in section 4 are part of the approximate inference

A: Yes, they are part of the inference procedure, the model remains the same as in section 3. These approximations are similar to mean-field approximations in variational inference.
\newline
\newline

Q: The assumption of mutual independence over the elements of S and beta_{l-1}

A: Generally speaking, correlation is partially removed due to the soft-thresholding on the previous layer. It requires further research to establish the influence of remaining correlation on the quality of approximation.
\newline
\newline

\section*{Reviewer 3}

\end{document}

### 2018-07-30 review_nips_2018/response_to_reviews_nips.tex

\section*{Reviewer 1}
Q: What is the motivation for being Bayesian about the LISTA algorithm. How does in the uncertainty in the Bayesian LISTA procedure relate to the original linear model.

A: The LISTA algorithm introduces the weights in the model. Their point estimates are being inferred during the training phase, this leads to the uncertainty of the estimation. In our Bayesian model we infer the posterior distributions over these weights to estimate this uncertainty.
\newline
\newline

Q: how does the sparse promoting prior interact with the sparsity promoting non-linearity.

A: In the LISTA model the sequential estimates of beta are obtained at each layer using the previous estimate. At every layer the estimate becomes more sparse due to soft-thresholding non-linearity. We do not specify sparsity priors over the $beta_l$ - we show that the distribution of these estimates can be approximated with the spike-and-slab distribution, i.e. its sparsity naturally arises from the Gaussian prior on NN weights and sparsity promoting soft thresholding nonlinearity.
\newline
\newline

@@

Q: The approximations in section 4 are part of the approximate inference

A: Yes, they are part of the inference procedure, the model remains the same as in section 3. These approximations are similar to mean-field approximations in variational inference.
\newline
\newline

Q: The assumption of mutual independence over the elements of $S$ and $beta_{l-1}$

### 2018-07-30 review_nips_2018/response_to_reviews_nips.tex

\section*{Reviewer 1}
Q: What is the motivation for being Bayesian about the LISTA algorithm. How does in the uncertainty in the Bayesian LISTA procedure relate to the original linear model.

A: The LISTA algorithm introduces the weights in the model. Their point estimates are being inferred during the training phase, this leads to the uncertainty of the estimation. In our Bayesian model we infer the posterior distributions over these weights to estimate this uncertainty.
\newline
\newline

Q: how does the sparse promoting prior interact with the sparsity promoting non-linearity.

A: In the LISTA model the sequential estimates of beta are obtained at each layer using the previous estimate. At every layer the estimate becomes more sparse due to soft-thresholding non-linearity. We do not specify sparsity priors over the $\beta_l$ - we show that the distribution of these estimates can be approximated with the spike-and-slab distribution, i.e. its sparsity naturally arises from the Gaussian prior on NN weights and sparsity promoting soft thresholding nonlinearity.
\newline
\newline

@@

Q: The approximations in section 4 are part of the approximate inference

A: Yes, they are part of the inference procedure, the model remains the same as in section 3. These approximations are similar to mean-field approximations in variational inference.
\newline
\newline

Q: The assumption of mutual independence over the elements of $S$ and $\beta_{l-1}$

### 2018-07-30 review_nips_2018/response_to_reviews_nips.tex

\documentclass{article}
\usepackage{nips_2018_author_response}
\usepackage{color}
\usepackage{mathtools}
\begin{document}

We would like to thank you the reviewers for their time and valuable suggestions.

\textbf{Reviewer 1}

\textbf{Q:} ``What is the motivation for being Bayesian about the LISTA algorithm. How does in the uncertainty in the Bayesian LISTA procedure relate to the original linear model.'' 
\textbf{A:} The LISTA algorithm introduces the weights in the model. Their point estimates are being inferred during the training phase, this leads to the uncertainty of the estimation. In our Bayesian model we infer the posterior distributions over these weights to estimate this uncertainty.

\textbf{Q:} ``how does the sparse promoting prior interact with the sparsity promoting non-linearity.'' 
\textbf{A:} In the LISTA model the sequential estimates of beta are obtained at each layer using the previous estimate. At every layer the estimate becomes more sparse due to soft-thresholding non-linearity. We do not specify sparsity priors over the $\beta_l$ - we show that the distribution of these estimates can be approximated with the spike-and-slab distribution, i.e. its sparsity naturally arises from the Gaussian prior on NN weights and sparsity promoting soft thresholding nonlinearity.

\textbf{Q:} ``The approximations in section 4 are part of the approximate inference.''  
\textbf{A:} Yes, they are part of the inference procedure, the model remains the same as in section 3. These approximations are similar to mean-field approximations in variational inference.

\textbf{Q:} ``The assumption of mutual independence over the elements of $S$ and $\beta_{l-1}$'' 
\textbf{A:} Generally speaking, correlation is partially removed due to the soft-thresholding on the previous layer. It requires further research to establish the influence of remaining correlation on the quality of approximation.

\textbf{Reviewer 2}

\textbf{Q:} ``the contribution is quite incremental, with most of the method relying on existing tools for probabilistic back-propagation'' 
\textbf{A:} The main contribution of the paper is in the forward pass, i.e. modelling distributions of the weights and how they propagate through the layers of a NN. We have shown that normally distributed weights propagated though the soft-thresholding non-linearity can be approximated with the spike and slab distribution. Although our inference is based on ideas from probabilistic back-propagation, we expanded this framework. In the original work, PBP was developed only for ReLU non-linearity, we developed it for soft-thresholding non-linearity. Also in contrast to the original work we deal with the multidimensional output of a NN and our NN is recurrent.

\textbf{Q:} ``Bayesian LISTA outperforms LISTA on NMSE (esp. with more layers/iterations), while LISTA yields better F- measures.''  
\textbf{A:} LISTA gets better F-measure only for number of measurements $K = \{10, 20\}$ for the synthetic data. In all the other experiments Bayesian LISTA shows either competitive results with LISTA or outperforms it.

\textbf{Q:} ``the MNIST experiment presented in the paper is about generating random Gaussian linear combinations of the MNIST images, and then trying to recover the original coefficients. No sparsity is imposed.'' 
\textbf{A:} Sparsity in the MNIST experiment is in images themselves. Flatten images represent vectors with a lot of zeros corresponding to background pixels.

\textbf{Reviewer 3}

\textbf{Q:} `` The update rules for the backpropagation algorithm should be given for each parameter, at least in the supplementary.'' 
\textbf{A:} We included details of the backpropagation algorithm that are distinguish and original for this paper. The general idea and formula of updating parameters based on derivatives of the normalisation constant is not new, therefore, we provided only derivation for the normalisation constant, which is different. We will add full details in the supplementary to make the paper self-contained. Thank you for your suggestion.

\textbf{Q:} ``My main concern is that the loss used to train the network is not detailed.'' 
\textbf{A:} Proposed Bayesian LISTA does not directly minimise any loss while training the network, we rather minimise the KL-divergence between approximating and posterior distribution on parameters. Backpropagation is used for the ADF updates of the parameters for computing derivatives of the normalisation constant. LISTA minimises the squared error between the predicted output and the given output for training network. Eq. (1) provides the general formulation of the sparse coding problem. It is not used for training networks. The parameter $\lambda$ was introduced in ISTA. The value $0.1$ was shown as this was found optimal for all algorithms on the synthetic data (plots are given in the supplementary).

\textbf{Q:} ``Only 200 samples are used from the 60,000 available, suggesting that the proposed technique scale poorly. The results are also displayed relatively to the number of training iteration and not with varying number L of layers/iterations as in the previous experiment'' 
\textbf{A:} A limited number of samples are used to show benefits of using a Bayesian method such as Bayesian LISTA over a frequentist method such as LISTA. As shown in the example of increasing the number of measurements when the data is not scarce, Bayesian LISTA and LISTA provides competitive results. The similar behaviour is demonstrated when we increase the size of training data. We will add the experiments with varying L as suggested. Thank you.

### 2018-08-01 review_nips_2018/response_to_reviews_nips.tex

\documentclass{article}
\usepackage{nips_2018_author_response}
\usepackage{color}
\usepackage{mathtools}
\begin{document}

We would like to thank the reviewers for their time and valuable suggestions.

@@

\textbf{Reviewer 1}

\textbf{Q:} \textit{What is the motivation for being Bayesian about the LISTA algorithm? How does in the uncertainty in the Bayesian LISTA procedure relate to the original linear model?}
\textbf{A:} The LISTA algorithm introduces the weights in the model. Their point estimates are being inferred during the training phase, this leads to the uncertainty of the estimation. In the proposed Bayesian model we infer the posterior distributions over these weights to estimate this uncertainty.

\textbf{Q:} \textit{How does the sparse promoting prior interact with the sparsity promoting non-linearity?}
\textbf{A:} In the LISTA model the sequential estimates of $\boldsymbol{\beta}$ are obtained at each layer using the previous estimate. At every layer the estimate becomes more sparse due to soft-thresholding non-linearity. We do not specify sparsity priors over the $\boldsymbol{\beta}_l$, we show that the distribution of these estimates can be approximated with the spike-and-slab distribution, i.e. its sparsity naturally arises from the Gaussian prior on NN weights and sparsity promoting soft thresholding nonlinearity.

\textbf{Q:} \textit{... whether the approximations in section 4 are part of the approximate inference} 
\textbf{A:} Yes, they are part of the inference procedure, the model remains the same as in section 3. These approximations are similar to mean-field approximations in variational inference.

\textbf{Q:} \textit{Is the assumption of mutual independence over the elements of $\mathbf{S}$ and $\boldsymbol{\beta}_{l-1}$?}
\textbf{A:} Generally speaking, correlation is partially removed due to the soft-thresholding on the previous layer. It requires further research to establish the influence of remaining correlation on the quality of approximation.

@@

\textbf{Reviewer 2}

\textbf{Q:} \textit{... the contribution is quite incremental, with most of the method relying on existing tools for probabilistic back-propagation}
\textbf{A:} The main contribution of the paper consists in the proposed way to model the distributions of the weights and their propagation through the layers of a NN. We have shown that normally distributed weights propagated though soft-thresholding non-linearity can be approximated with the spike and slab distribution. Although our inference is based on ideas from probabilistic backpropagation, we expanded this framework. In the original work, PBP was developed only for ReLU non-linearity, we developed it for soft-thresholding non-linearity. Also in contrast to the original work we dealt with the multidimensional output of a NN and considered NNs are recurrent.

\textbf{Q:} \textit{Bayesian LISTA outperforms LISTA on NMSE (esp. with more layers/iterations), while LISTA yields better F- measures} 
\textbf{A:} LISTA gets better F-measure only for the number of measurements $K = \{10, 20\}$ for the synthetic data. In all the other experiments Bayesian LISTA shows either competitive results with LISTA or outperforms it.

\textbf{Q:} \textit{... the MNIST experiment presented in the paper is about generating random Gaussian linear combinations of the MNIST images, and then trying to recover the original coefficients. No sparsity is imposed}
\textbf{A:} Sparsity in the MNIST experiment is in images themselves. Flatten images represent vectors with a lot of zeros corresponding to background pixels.

@@

\textbf{Reviewer 3}

\textbf{Q:} \textit{The update rules for the backpropagation algorithm should be given for each parameter, at least in the supplementary} 
\textbf{A:} We included details of the backpropagation algorithm that are original for this paper. The general idea and formula of updating parameters based on derivatives of the normalisation constant is not new, therefore, we provided only derivation for the normalisation constant, which is different. We will add full details in the supplementary to make the paper self-contained. Thank you for your suggestion.

\textbf{Q:} \textit{My main concern is that the loss used to train the network is not detailed}
\textbf{A:} Proposed Bayesian LISTA does not directly minimise any loss while training the network, we rather minimise the KL-divergence between approximating and posterior distribution on parameters. Backpropagation is used for the ADF updates of the parameters for computing derivatives of the normalisation constant. LISTA minimises the squared error between the predicted output and the given output for training network. Eq. (1) provides the general formulation of the sparse coding problem. It is not used for training networks. The parameter $\lambda$ was introduced in ISTA. The value $\lambda=0.1$ was shown to be optimal for all algorithms on the synthetic data (plots are given in the supplementary).

\textbf{Q:} \textit{Only 200 samples are used from the 60,000 available, suggesting that the proposed technique scale poorly. The results are also displayed relatively to the number of training iteration and not with varying number $L$ of layers/iterations as in the previous experiment...} 
\textbf{A:} A limited number of samples are used to show benefits of using a Bayesian method such as Bayesian LISTA over a frequentist method such as LISTA. As shown in the example of increasing the number of measurements when the data is not scarce, Bayesian LISTA and LISTA provide competitive results in terms of accuracy of point estimates. The similar behaviour is demonstrated when we increase the size of training data. We will add the experiments with varying $L$ as suggested. Thank you.

### 2018-10-23 icassp_2019/paper.tex
started icassp version

\documentclass{article}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{xcolor}

\usepackage{times}
\usepackage{float}

\usepackage[square,numbers]{natbib}
\usepackage[nonatbib]{nips_2018}

\usepackage{bbm}
\usepackage{algorithm, algorithmic}

@@

\allowdisplaybreaks

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicreturn}{\textbf{Output:}}

% \setlength{\belowcaptionskip}{-15pt}
%  \usepackage[small,compact]{titlesec}
%  \titlespacing{\section}{0pt}{1ex}{0.8ex}
%  \titlespacing{\subsection}{0pt}{0.5ex}{0ex}
%  \titlespacing{\subsubsection}{0pt}{0.2ex}{0ex}
%  \expandafter\def\expandafter\normalsize\expandafter{%
%      \normalsize
%      \setlength\abovedisplayskip{2pt}
%      \setlength\belowdisplayskip{2pt}
%      \setlength\abovedisplayshortskip{0pt}
%      \setlength\belowdisplayshortskip{0pt}
%  }
%  \renewcommand{\baselinestretch}{0.99}

\title{Uncertainty propagation in neural networks for sparse coding}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

%\author{ {\bf Danil Kuzin} \\
%\And
%{\bf Olga Isupova}  \\
%\And
%{\bf Lyudmila Mihaylova}   \\
%}

@@

\section{Introduction}
The idea of Bayesian learning in neural networks \cite{neal2012bayesian} has recently regained its attention with the development of distributed approximate inference techniques \cite{li2015stochastic, hoffman2013stochastic} and general boost in popularity of deep learning. In addition to algorithms' predictions themselves, Bayesian learning naturally provides the uncertainty estimates of these predictions. These uncertainty estimates are vital, e.g., in spheres that affect people's health, such as self-driving cars or medicine.

@@

%The paper is organised as following: first, we review neural networks for sparse coding in section \ref{sec:nn_sc} and present a novel Bayesian neural network in section \ref{sec:bayesian_lista}. We then describe uncertainty propagation to obtain the spike and slab distribution for an output of a network in section \ref{sec:fprop} and how probabilistic backpropagation is organised to update distributions of weights in section \ref{sec:backpropagation}. After that, we show the experimental results of the algorithm in section \ref{sec:experiments} and finish with discussion and conclusions in section \ref{sec:conclusions}.

@@

% \begin{align}
% \label{eq:first_layer}
% &\mathbf{b} = \mathbf{W}\mathbf{y}\\
% \label{eq:thr_first}
% &\widehat{\boldsymbol\beta}_0 = h_\lambda(\mathbf{b}) \\
% &\text{for } l=1:L\\
% \label{eq:l_dense_layer}
% 	&\quad \mathbf{c}_l = \mathbf{b} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1} \\
% \label{eq:l_thr}
% 	&\quad \widehat{\boldsymbol\beta}_{l} = h_\lambda(\mathbf{c}_l) \\
% & \widehat{\boldsymbol\beta} = \widehat{\boldsymbol\beta}_{L}
% \end{align}

@@

%The shrinkage parameter $\lambda$ is currently treated as a deterministic hyperparameter for Bayesian \textsc{lista}. In future we plan to incorporate it into the model treating it as a random variable. For this we will extend both the uncertainty propagation method to include uncertainty of $\lambda$ and the probabilistic backpropagation algorithm to estimate its posterior.

\small
\bibliography{bibliography}
\bibliographystyle{unsrtnat}

>>>>

\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{times}
\usepackage{float}
\usepackage{algorithm, algorithmic}

@@

\title{Bayesian neural networks for sparse coding}

\name{Danil Kuzin$^{\dagger}$ \qquad Olga Isupova$^{\star}$ \qquad Lyudmila Mihaylova$^{\dagger}$}

\address{$^{\star}$ Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield S1 3JD, UK (e-mail: dkuzin1@sheffield.ac.uk; l.s.mihaylova@sheffield.ac.uk).\\
  		    $^{\dagger}$ Department of Engineering Science, University of Oxford, Oxford OX1 2JD, UK (e-mail: olga.isupova@eng.ox.ac.uk)}

@@

\begin{keywords}
Sparse regression, Bayesian neural networks, uncertainty estimation, compressive sensing
\end{keywords}

\section{Introduction}
\label{sec:intro}

  The idea of Bayesian learning in neural networks \cite{neal2012bayesian} has recently regained its attention with the development of distributed approximate inference techniques \cite{li2015stochastic, hoffman2013stochastic} and general boost in popularity of deep learning. In addition to algorithms' predictions themselves, Bayesian learning naturally provides the uncertainty estimates of these predictions. These uncertainty estimates are vital, e.g., in spheres that affect people's health, such as self-driving cars or medicine.

@@

@@

@@

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
% \begin{figure}[htb]

% \begin{minipage}[b]{1.0\linewidth}
%   \centering
%   \centerline{\includegraphics[width=8.5cm]{image1}}
% %  \vspace{2.0cm}
%   \centerline{(a) Result 1}\medskip
% \end{minipage}
% %
% \begin{minipage}[b]{.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image3}}
% %  \vspace{1.5cm}
%   \centerline{(b) Results 3}\medskip
% \end{minipage}
% \hfill
% \begin{minipage}[b]{0.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image4}}
% %  \vspace{1.5cm}
%   \centerline{(c) Result 4}\medskip
% \end{minipage}
% %
% \caption{Example of placing a figure with experimental results.}
% \label{fig:res}
% %
% \end{figure}

% % To start a new column (but not a new page) and help balance the last-page
% % column length use \vfill\pagebreak.
% % -------------------------------------------------------------------------
% %\vfill
% %\pagebreak

% \section{COPYRIGHT FORMS}
% \label{sec:copyright}

% You must submit your fully completed, signed IEEE electronic copyright release
% form when you submit your paper. We {\bf must} have this form before your paper
% can be published in the proceedings.

% \section{RELATION TO PRIOR WORK}
% \label{sec:prior}

% The text of the paper should contain discussions on how the paper's
% contributions are related to prior work in the field. It is important
% to put new work in  context, to give credit to foundational work, and
% to provide details associated with the previous work that have appeared
% in the literature. This discussion may be a separate, numbered section
% or it may appear elsewhere in the body of the manuscript, but it must
% be present.

% You should differentiate what is new and how your work expands on
% or takes a different path from the prior studies. An example might
% read something to the effect: "The work presented here has focused
% on the formulation of the ABC algorithm, which takes advantage of
% non-uniform time-frequency domain analysis of data. The work by
% Smith and Cohen \cite{Lamp86} considers only fixed time-domain analysis and
% the work by Jones et al \cite{C2} takes a different approach based on
% fixed frequency partitioning. While the present study is related
% to recent approaches in time-frequency analysis [3-5], it capitalizes
% on a new feature space, which was not considered in these earlier
% studies."

% \vfill\pagebreak

% \section{REFERENCES}
% \label{sec:refs}

% List and number all bibliographical references at the end of the
% paper. The references can be numbered in alphabetic order or in
% order of appearance in the document. When referring to them in
% the text, type the corresponding reference number in square
% brackets as shown at the end of this sentence \cite{C2}. An
% additional final page (the fifth page, in most cases) is
% allowed, but must contain only references to the prior
% literature.

% % References should be produced using the bibtex program from suitable
% % BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% % style file from IEEE produces unsorted bibliography list.
% % -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{bibliography}
